[![Travis build status](https://travis-ci.org/dpastoor/rbabylon.svg?branch=master)](https://travis-ci.org/dpastoor/rbabylon)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/dpastoor/rbabylon?branch=master&svg=true)](https://ci.appveyor.com/project/dpastoor/rbabylon)
 
rbabylon
============

An R interface to the babylon ecosystem.

## Babylon queue (bbq) interface

`bbq` is a nonmem queue and model runner that exposes an http API for model submission and management.

The overall structure of the information kept in bbq is a tree-like, therefore represented in R as a list.

The primary fields are:

* id
* status
* model_info
    * model_path
    * run_settings
        * git
        * save_exe
        * verbose
        * debug
        * clean_lvl
        * copy_lvl
        * cache_dir
        * exe_name_in_cache
        * nm_executable_or_path
        * one_est
        * proposed_run_dir
* run_info
    * queue_time,
    * start_time,
    * duration,
    * run_dir,
    * error


For example:

```json
{
	"ID": 11,
	"Status": "COMPLETED",
	"ModelInfo": {
		"ModelPath": "C:\\golang\\src\\github.com\\dpastoor\\babylon\\testdata\\run001.mod",
		"RunSettings": {
			"Git": false,
			"SaveExe": "",
			"Verbose": false,
			"Debug": false,
			"CleanLvl": 5,
			"CopyLvl": 5,
			"CacheDir": "C:\\golang\\src\\github.com\\dpastoor\\babylon\\testdata\\mdlcache",
			"ExeNameInCache": "run001.exe",
			"NmExecutableOrPath": "nmfe74",
			"OneEst": false,
			"ProposedRunDir": ""
		}
	},
	"RunInfo": {
		"QueueTime": 0,
		"StartTime": 1500764944,
		"Duration": 0,
		"RunDir": "C:\\golang\\src\\github.com\\dpastoor\\babylon\\testdata\\run001_est_07",
		"Error": ""
	}
}
```

## rbabylon

Interaction with bbq in R can occur over the http api to submit and query models. An ORM of sorts
is exposed through `rbabylon` as the Babylon class:

```{r eval = F}
bbq <- bbq_client()
# or 
bbq <- Babylon$new()
```

### submitting models

model submission is done through the `submit_models` method. The only required bit of information
is a path to the model file(s) to submit. If the path is not absolute, it will be considered relative
to the current working directory. One mechanism is to call `normalizePath()` on the paths to let
R clean and absolutify a path. For example, to submit all models in the dir "modeling/"
```{r eval = F}
models <- normalizePath(dir("modeling", pattern = "*.mod", full.names = TRUE))
response <- bbq$submit_models(models)
```

The returned response object will contain the model IDs set by the server, so the submitted models can be
easily queried

```{r eval = F}
model_ids <- reponse %>% map_dbl("id")
```


### getting model information

To get all models, can query with `bbq$get_models()`, which will
provide a list of lists (a list of model-information-lists) for each run.

For example, to extract a vector run directories for all complete models one could do:

```
bbq$get_models(status = "COMPLETED") %>% map_chr(~ .x$RunInfo$RunDir)
```

Hence, it is quite valuable to learn some functional programming techniques made available through
the [purrr](http://purrr.tidyverse.org/) package


### polling

One common need is to run model(s) and do something after they are completed. If you are willing/want to block
your R session, such that the remainder of code will not execute until the submitted models are completed
you may use the `poll()` method, which takes a vector of model ID's, and will query the bbq server at
intervals specified by the `interval` argument (default 10 seconds) until all models have one of the statuses
specified in the `until` argument (default `COMPLETED` and `ERROR`).

After all models achieve the specified status(es), or the `timeout` time has elapsed, the model's information
will be returned, or if the timeout is elapsed, a NULL result is returned.

This allows programmatic flows such as:

```{r eval=F}
bbq$poll(c(1, 2, 3)) %>%
    # do postprocessing
    post_process_models() %>%
    ...etc
```

The 'downside' is it blocks your R-session while polling, though this behavior is sometimes desired, especially
for short running models/simulations.

Regardless, to asyncronously monitor run status, other mechanisms such as `slack` or text notifictions, 
or monitoring through a separate web GUI are likely more ideal.
